The Astropy Project: Building an inclusive, open science project and status of the v2.0 software

### 2018-01-08 main.tex
added link to repository with paper
added links to CI section and edits

\begin{abstract}
% I (Adrian) took a first stab at the abstract, but it needs some work. Feel
% free to modify!
The \astropy project supports and fosters the development of open-source and openly-developed
\python packages that provide commonly-needed functionality to the astronomical
community.
A key element of the \astropy project is the core package \astropypkg, which serves as the
foundation for more specialized projects and packages.
In this article, we provide an overview of the organization of the \astropy
project and summarize key features in the core package as of the recent major
release, version 2.0.
We then describe the project infrastructure designed to facilitate and support
development for a broader ecosystem of inter-operable packages.
We conclude with a future outlook of planned new features and directions for the
broader \astropy project.
\end{abstract}

% Adrian PW: how are there still no software keywords in AAS journals?!?
\keywords{%
    Astrophysics - Instrumentation and Methods for Astrophysics
    ---
    methods: data analysis
    ---
    methods: miscellaneous
}

@@

We start by describing the way the \astropy project functions and is organized
in \sectionname~\ref{sec:org}. We then describe the main software efforts
developed by the \astropy project itself: a core package called \astropypkg
(\sectionname~\ref{sec:core}) and several separate packages that help maintain
the infrastructure for testing and documentation
(\sectionname~\ref{sec:infrastructure}). We end with a short vision for
the future of \astropy and astronomical software in general in
\sectionname~\ref{sec:future}.

@@

\subsection{Coordination of Astropy}
% draft by ?? (BMS thinks Erik)
% edited by Adrian PW
\label{sect:coordcom}
From its inception, \astropy has required coordination to ensure the project
as a whole and its coding efforts are consistent and reasonably efficient.
While many \python projects adopt a ``Benevolent Dictator For Life'' (BDFL)
model, \astropy has instead opted for a \emph{coordination committee}.  This
is in part due to the nature of the project as a large-scale collaboration
between many contributors with many interests, and in part due to simply the
amount of work that needs to get done.  For the latter reason, the
project has expanded the committee from three to four members starting in
2016.

@@

\subsection{Astropy development model}
% draft by Adrian PW and Erik T.
Code is contributed to the \astropypkg core package or modified through ``pull
requests'' (via \github) that often contain several \texttt{git} commits.
Pull requests may fix bugs, implement new features, or improve or modify the
infrastructure that supports the development and maintenance of the package.
Individual pull requests are generally limited to a single conceptual addition
or modification to make code review tractable.
Pull requests that modify or add code to a specific subpackage must be reviewed
and approved by one of the subpackage maintainers before they are merged into
the core codebase.
Bugs and feature requests are reported via the \github issue tracker and labeled
with a set of possible labels that help classify and organize the issues.
The development workflow is detailed in the \astropypkg
documentation.\footnote{\emph{How to make a code contribution},
\url{http://docs.astropy.org/en/latest/development/workflow/development_workflow.html}}

@@

\subsection{APEs - Astropy Proposals for Enhancement}
% draft by hamogu
% edited by Adrian PW

@@

\subsection{Concept of affiliated packages}
% it may be described in previous subsection, if not the Brigitta can try to
% write it up
% edited by Adrian PW

@@

Affiliated packages contain functionality that is more specialized,
have license incompatibilities, or have external dependencies (e.g., GUI
libraries) that make these packages more suitable to be separate from the
\astropypkg core package.
Affiliated packages may also be used to develop substantial new functionality
that will eventually be incorporated into the \astropypkg core package
(e.g., \texttt{wcsaxes}).
New functionality benefits from having a rapid development and release cycle that is not tied to that of the \astropypkg core (\sectionname~\ref{sect:releasecycle}).
% AG: rephrased this.
%Initially separate packages with typically rapid initial development and new experimental features benefit from having a release cycle that is not tied to the

@@

%\subsection{Accuracy testing across many different implementation}

\subsection{Release cycle and Long Term Support}
\label{sect:releasecycle}
% This could also include information about the development cycle
% Brigitta can writes this up if there's no other taker
% edited by Adrian PW

@@

The released versions of the \astropypkg core package are available from several
of the \python distributions for scientific computing (e.g.,
\href{http://anaconda.org}{Anaconda}) and from the \python Package Index
(PyPI).\footnote{See the installation documentation for more information:
\url{http://docs.astropy.org/en/stable/install.html}}
Effort has been made to make \astropypkg available and easily installable across
all platforms; the package is constantly tested on different platforms as part
of a suite of continuous integration (CI) tests.

@@

\section{Astropy Core Package Version 2.0}
\label{sec:core}
% Adrian
The \astropy project aims to provide \python-based packages for all tasks that
are commonly needed in a large subset of the astronomical community.
At the foundation is the \astropypkg core package, which provides general
functionality (e.g., coordinate transformations, reading and writing
astronomical files, and units) or base classes for other
packages to utilize for a common interface (e.g., \texttt{NDData}).
In this section, we highlight new features introduced or substantially improved
since version 0.2 (previously described in \citealt{astropy}).  The \astropypkg
package
provides a full log of changes\footnote{\url{https://github.com/astropy/astropy/blob/stable/CHANGES.rst}}
over the course of the entire project and more details about individual
subpackages are available in the documentation.\footnote{\url{http://docs.astropy.org/en/stable/}}
Beyond what is mentioned below, most subpackages have seen improved performance
since the release of the version 0.2 package.

\subsection{Units}\label{sec:units}
% Adrian

@@

\subsection{Constants}
% David S.
% Edited by Adrian PW

@@

\subsection{Coordinates}
\label{sec:coordinates}
% Adrian, Erik
The \package{astropy.coordinates} subpackage is designed to support representing
and transforming celestial coordinates and, new in version 2.0, velocities.
The framework heavily relies on the \package{astropy.units} subpackage, and most
inputs to objects in this subpackage are expected to be \texttt{Quantity}
objects.
Some of the machinery also relies on the Essential Routines of Fundamental
Astronomy (ERFA) \texttt{C} library for some of the critical underlying
transformation machinery \citep{erfa}, which is based on the Standards Of
Fundamental Astronomy (SOFA) effort \citep{sofa}.

@@

\subsection{Time}
\label{sec:time}
% Adrian PW

@@

\subsubsection{Tables}
\label{sec:table}
% fleshed out / edited by Adrian PW

@@

\subsection{\package{io}}
% Simon

@@

\subsection{Modeling}
\label{sec:modeling}
% The whole modeling submodule was missing from the previous paper, so everything really, including compound models, unit support etc.
% Nadia, Lim
% Tom R. -- unit support

@@

\subsection{Convolution}
% Adam G.

@@

\subsection{Visualization}
% Larry:  Image visualization (stretching, scaling), RGB

@@

%\subsection{Utils}
% Lim

@@

\subsection{Statistics}
% Edited by Adrian PW

@@

% Steve C.: overview, circular stats,
% Jake V.:  Lomb-scargle, Bayesian blocks
% Larry:  sigma clipping, biweight stats
% Ze':  Ripley's K (spatial stats)

@@

\label{sec:infrastructure}
% Text also edited by Adrian PW

@@

\subsection{Package template}
% Brigitta may write this up if there is no other takers

@@

\subsection{Continuous integration helpers}
% Brigitta

\astropy also provides a set of scripts for setting up and configuring
CI services as a \github repository,
\package{astropy/ci-helpers}.\footnote{\url{https://github.com/astropy/ci-helpers}}
These tools aim to enable package maintainers to control their testing setup
and installation process for various CI services through a
set of environment variables.
While the current development is mostly driven by the needs of the \astropy
ecosystem, the actual usage of this package is extremely widespread. The current
tools support configuration for Travis CI and Appveyor CI.

@@

\section{The future of the Astropy project}
\label{sec:future}
% First draft by SMC
% edited by BMS
% edited again by Adrian PW

@@

\subsection{Learn Astropy}
% Adrian

@@

\section{Conclusion}
\label{sec:conclusion}
% Renamed and edited by Adrian PW

@@

We would like to thank the members of the community that have 
contributed to \astropy, 
that have opened issues and provided feedback, and have supported the 
project in a number of different ways.  We would like to acknowledge
Alex Conley and Neil Crighton for maintaining the  \package{cosmology} 
subpackage. 
%insert significant contributors to astropy who have decided not to 
%be co-authors

>>>>

\begin{abstract}
The \astropy project supports and fosters the development of open-source and openly-developed
\python packages that provide commonly-needed functionality to the astronomical
community.
A key element of the \astropy project is the core package \astropypkg, which serves as the
foundation for more specialized projects and packages.
In this article, we provide an overview of the organization of the \astropy
project and summarize key features in the core package as of the recent major
release, version 2.0.
We then describe the project infrastructure designed to facilitate and support
development for a broader ecosystem of inter-operable packages.
We conclude with a future outlook of planned new features and directions for the
broader \astropy project.
\end{abstract}

\keywords{%
    Astrophysics - Instrumentation and Methods for Astrophysics
    ---
    methods: data analysis
    ---
    methods: miscellaneous
}

@@

We start by describing the way the \astropy project functions and is organized
in \sectionname~\ref{sec:org}. We then describe the main software efforts
developed by the \astropy project itself: a core package called \astropypkg
(\sectionname~\ref{sec:core}) and several separate packages that help maintain
the infrastructure for testing and documentation
(\sectionname~\ref{sec:infrastructure}). We end with a short vision for
the future of \astropy and astronomical software in general in
\sectionname~\ref{sec:future}. The full paper, including the code
to produce the figures, is available in a \github 
repository\footnote{\url{https://github.com/astropy/astropy-v2.0-paper}}

@@

\subsection{Coordination of Astropy}
\label{sect:coordcom}
From its inception, \astropy has required coordination to ensure the project
as a whole and its coding efforts are consistent and reasonably efficient.
While many \python projects adopt a ``Benevolent Dictator For Life'' (BDFL)
model, \astropy has instead opted for a \emph{coordination committee}.  This
is in part due to the nature of the project as a large-scale collaboration
between many contributors with many interests, and in part due to simply the
amount of work that needs to get done.  For the latter reason, the
project has expanded the committee from three to four members starting in
2016.

@@

\subsection{Astropy development model}

Code is contributed to the \astropypkg core package or modified through ``pull
requests'' (via \github\footnote{\url{https://github.com/astropy/astropy/}}) 
that often contain several \texttt{git} commits.
Pull requests may fix bugs, implement new features, or improve or modify the
infrastructure that supports the development and maintenance of the package.
Individual pull requests are generally limited to a single conceptual addition
or modification to make code review tractable.
Pull requests that modify or add code to a specific subpackage must be reviewed
and approved by one of the subpackage maintainers before they are merged into
the core codebase.
Bugs and feature requests are reported via the \github issue tracker and labeled
with a set of possible labels that help classify and organize the issues.
The development workflow is detailed in the \astropypkg
documentation.\footnote{\emph{How to make a code contribution},
\url{http://docs.astropy.org/en/latest/development/workflow/development_workflow.html}}

@@

\subsection{APEs - Astropy Proposals for Enhancement}

@@

\subsection{Concept of affiliated packages}

@@

Affiliated packages contain functionality that is more specialized,
have license incompatibilities, or have external dependencies (e.g., GUI
libraries) that make these packages more suitable to be separate from the
\astropypkg core package.
Affiliated packages may also be used to develop substantial new functionality
that will eventually be incorporated into the \astropypkg core package
(e.g., \texttt{wcsaxes}).
New functionality benefits from having a rapid development and release cycle that is not tied to that of the \astropypkg core (\sectionname~\ref{sect:releasecycle}).

@@

\subsection{Release cycle and Long Term Support}
\label{sect:releasecycle}

@@

The released versions of the \astropypkg core package are available from several
of the \python distributions for scientific computing (e.g.,
\href{http://anaconda.org}{Anaconda}) and from the \python Package Index
(PyPI).\footnote{See the installation documentation for more information:
\url{http://docs.astropy.org/en/stable/install.html}}
Effort has been made to make \astropypkg available and easily installable across
all platforms; the package is constantly tested on different platforms as part
of a suite of continuous integration tests.

@@

\section{Astropy Core Package Version 2.0}
\label{sec:core}
The \astropy project aims to provide \python-based packages for all tasks that
are commonly needed in a large subset of the astronomical community.
At the foundation is the \astropypkg core package, which provides general
functionality (e.g., coordinate transformations, reading and writing
astronomical files, and units) or base classes for other
packages to utilize for a common interface (e.g., \texttt{NDData}).
In this section, we highlight new features introduced or substantially improved
since version 0.2 (previously described in \citealt{astropy}).  The \astropypkg
package
provides a full log of changes\footnote{\url{https://github.com/astropy/astropy/blob/stable/CHANGES.rst}}
over the course of the entire project and more details about individual
subpackages are available in the documentation.\footnote{\url{http://docs.astropy.org/en/stable/}}
Beyond what is mentioned below, most subpackages have seen improved performance
since the release of the version 0.2 package.

\subsection{Units}\label{sec:units}

@@

\subsection{Constants}

@@

\subsection{Coordinates}
\label{sec:coordinates}
The \package{astropy.coordinates} subpackage is designed to support representing
and transforming celestial coordinates and, new in version 2.0, velocities.
The framework heavily relies on the \package{astropy.units} subpackage, and most
inputs to objects in this subpackage are expected to be \texttt{Quantity}
objects.
Some of the machinery also relies on the Essential Routines of Fundamental
Astronomy (ERFA) \texttt{C} library for some of the critical underlying
transformation machinery \citep{erfa}, which is based on the Standards Of
Fundamental Astronomy (SOFA) effort \citep{sofa}.

@@

\subsection{Time}
\label{sec:time}

@@

\subsubsection{Tables}
\label{sec:table}

@@

\subsection{\package{io}}

@@

\subsection{Modeling}
\label{sec:modeling}

@@

\subsection{Convolution}

@@

\subsection{Visualization}

@@

@@

\subsection{Statistics}

@@

@@

\label{sec:infrastructure}

@@

\subsection{Package template}

@@

\subsection{Continuous integration helpers}

\astropy also provides a set of scripts for setting up and configuring
continuous integration (CI) services as a \github repository,
\package{astropy/ci-helpers}.\footnote{\url{https://github.com/astropy/ci-helpers}}
These tools aim to enable package maintainers to control their testing setup
and installation process for various CI services through a
set of environment variables.
While the current development is mostly driven by the needs of the \astropy
ecosystem, the actual usage of this package is extremely widespread. The current
tools support configuration for Travis CI\footnote{\url{https://travis-ci.org/}} and 
Appveyor CI\footnote{\url{https://www.appveyor.com/}}.

@@

\section{The future of the Astropy project}
\label{sec:future}

@@

\subsection{Learn Astropy}

@@

\section{Conclusion}
\label{sec:conclusion}

@@

We would like to thank the members of the community that have 
contributed to \astropy, 
that have opened issues and provided feedback, and have supported the 
project in a number of different ways.  We would like to acknowledge
Alex Conley and Neil Crighton for maintaining the  \package{cosmology} 
subpackage.
